# Reproducible Research: Peer Assessment 1

## Loading and preprocessing the data

First, unzip the data file and read it into R. Add a new column which converts the interval from military time to minutes since midnight.

```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv", colClasses = c("integer", "Date", "integer"))
activity$intervalminutes <- activity$interval %/% 100 * 60 + activity$interval %% 100
```

## What is mean total number of steps taken per day?

Calculate total steps per day, plot a histogram and report mean and median

```{r hist}
dailysteps <- aggregate(steps~date, data = activity, FUN = "sum")
hist(dailysteps$steps, main = "Histogram of Total Steps per Day", xlab = "Total Steps per Day")
```

Mean steps per day:
```{r}
mean(dailysteps$steps)
```

Median steps per day:
```{r}
median(dailysteps$steps)
```

## What is the average daily activity pattern?

Calculate average for each interval, plot daily activity pattern and report interval with maximum number of steps.
Note: Plot uses minutes past midnight instead of original intervals (military time) in order to prevent "jumps" at the hour boundaries.

```{r daily}
avgsteps <- aggregate(steps~intervalminutes + interval, data = activity, FUN = "mean")
plot(avgsteps$intervalminutes, avgsteps$steps, type = "l", main = "Average Daily Activity", xlab = "Minutes", 
     ylab = "Average Number of Steps")
```

Interval with maximum number of steps:

```{r}
avgsteps[which.max(avgsteps$steps),]
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
